...

### 2024-12-04 paper/paper.qmd
finished results section, added a new folder in "plan" with maps, and rendered pdf

For the present analysis, the variables of interest are the different types of census family sizes along with the average after-tax income for each of Toronto’s neighbourhoods. The Neighbourhood Profiles data includes four different census family sizes: (1) 2-person families, (2) 3-person families, (3) 4-person families, and (4) five or more-person families. In simple terms, Statistics Canada defines a census family as one where all family members (related by blood marriage, common-law union, adoption, or a foster relationship) live together in the same dwelling [@censusdefinition]. Census families can also be referred to as economic families [@economicdefinition]. The data further provides the average after-tax income ($) recorded in 2020 for each neighbourhood.

@@

@@

The construction of a model for this analysis began by fitting an initial model with average neighbourhood income as the response variable and 2, 3, 4, and 5 or more-person families as predictors. As seen by @tbl-initialmodel and the individual t-tests, 5 or more-person families do not appear to significantly impact a neighbourhood’s average income with the largest p-value among all predictors (p > 0.001).

@@

So, another model was created without this variable (i.e. a reduced model), and all predictors seem to have significant impacts (p < 0.001) on neighbourhood income as shown in @tbl-newmodel.

@@

par(mfrow=c(1,3))
plot(residuals_1 ~ regression_data$size_two, main = "Residuals vs 2-person Families", 
     xlab = "Number of 2-person Families", ylab = "Residuals")
plot(residuals_1 ~ regression_data$size_three, main = "Residuals vs 3-person Families", 
     xlab = "Number of 3-person Families", ylab = "Residuals")
plot(residuals_1 ~ regression_data$size_four, main = "Residuals vs 4-person Families", 
     xlab = "Number of 4-person Families", ylab = "Residuals")
```
The Residual vs Fitted Values plot (@fig-residualsandfitted) seems to have some clustering of points, indicating that the constant variance and independence assumptions are somewhat satisfied. However, because the overall pattern of the plot is still relatively spread out with no observable patterns, and as the objective of the model is interpretability, transformations were not applied to the variables in the model to better these assumptions.

@@

plot(residuals_1 ~ fitted_1, main = "Residuals vs Fitted Values", 
     xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col ="red")
```
Lastly, though the Normal Q-Q plot (@fig-normal) shows some deviation from the Q-Q line, the normality assumption is somewhat satisfied with little evidence to suggest an alarming violation of this assumption. Putting all this together, while improvements can be made to the model for better, more accurate predictions, it is sufficient for the purpose of interpretability.

>>>>

For the present analysis, the variables of interest are the different types of census family sizes along with the average after-tax income for each of Toronto’s neighbourhoods. The Neighbourhood Profiles data includes four different census family sizes: (1) 2-person families, (2) 3-person families, (3) 4-person families, and (4) five or more-person families. In simple terms, Statistics Canada defines a census family as one where all family members (related by blood marriage, common-law union, adoption, or a foster relationship) live together in the same dwelling [@censusdefinition]. Census families can also be referred to as economic families [@economicdefinition]. The data further provides the average after-tax income (\$) recorded in 2020 for each neighbourhood.

@@

## Distribution of Different Census Family Sizes Across Toronto Neighbourhoods

```{r}
#| echo: false 
#| message: false
#| label: fig-familysizetwo
#| fig-cap: Number of 2-person Census Families Across Toronto Neighbourhoods

family_size <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
#view(family_size)

# Rename columns

fam_size_data <-
  family_size |>
  rename(
    neighbourhood_name = "Neighbourhood Name",
    neighbourhood_num = "Neighbourhood Number",
    num_census_fam = "Total - Census families in private households by family size - 25% sample data",
    size_two = "2 persons",
    size_three = "3 persons",
    size_four = "4 persons",
    size_five_plus = "5 or more persons",
    avg_fam_size = "Average size of census families",
    avg_num_child = "Average number of children in census families with children",
    avg_income = "Average after-tax income of economic family in 2020 ($)"
  )

fam_size_data$size_two = as.numeric(as.character(fam_size_data$size_two))
fam_size_data$size_three = as.numeric(as.character(fam_size_data$size_three))
fam_size_data$size_four = as.numeric(as.character(fam_size_data$size_four))

ggplot(fam_size_data, aes(x = neighbourhood_num, y = size_two)) + 
  geom_bar(stat = "identity", colour = "blue", fill = "lightblue") + theme_light() +  
  ggtitle("2-person Families in Toronto Neighbourhoods") +
  labs(x = "158 Toronto Neighbourhoods", y = "Number of 2-person Families", 
       caption = "Data Source: Statistics Canada, 2021 Census") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_blank()
  )
```

```{r}
#| echo: false 
#| message: false
#| label: fig-familysizethree
#| fig-cap: Number of 3-person Census Families Across Toronto Neighbourhoods

ggplot(fam_size_data, aes(x = neighbourhood_num, y = size_three)) + 
  geom_bar(stat = "identity", colour = "purple", fill = "pink") + theme_light() +  
  ggtitle("3-person Families in Toronto Neighbourhoods") +
  labs(x = "158 Toronto Neighbourhoods", y = "Number of 3-person Families", 
       caption = "Data Source: Statistics Canada, 2021 Census") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_blank()
  )
```

```{r}
#| echo: false 
#| message: false
#| label: fig-familysizefour
#| fig-cap: Number of 4-person Census Families Across Toronto Neighbourhoods

ggplot(fam_size_data, aes(x = neighbourhood_num, y = size_four)) + 
  geom_bar(stat = "identity", colour = "darkgreen", fill = "lightgreen") + theme_light() +  
  ggtitle("4-person Families in Toronto Neigbourhoods") +
  labs(x = "158 Toronto Neighbourhoods", y = "Number of 4-person Families", 
       caption = "Data Source: Statistics Canada, 2021 Census") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_blank()
  )
```

@fig-familysizetwo, @fig-familysizethree, and @fig-familysizefour illustrate the number of 2, 3, and 4-person families, respectively, across Toronto’s 158 neighbourhoods. From this, 2-person census families appear to be the most common as almost all neighbourhoods have close to 1000 2-person families with several of them having over 3000, near 4000 2-person families. On the other hand, 3 and 4-person families seem to less common with more neighbourhoods having less than 1000 families of these sizes (i.e. under or just over 500 families) and with the maximum number of 3 or 4-person families being just under 3000. These observations seem to suggest that given the higher numbers of 2-person families across Toronto neighbourhoods, this family size may be a key driver of average neighbourhood income within the city of Toronto.

## Model Results

```{r}
#| warning: false
#| echo: false
#| label: tbl-modelresults
#| tbl-cap: "Summary Results of Analysis Model"

model_results <- readRDS(here("model/model_1.rds"))

modelsummary(model_results, statistic = "{std.error} ({p.value})")
```

The results from the model shown in @tbl-modelresults indicate significant relationships between average neighbourhood income and different census family sizes (p \< 0.001 for each family size). This means that the addition of a 2-person census family to a neighbourhood is expected to increase the neighbourhood’s average after-tax income by around \$28.59 (± 4.76). Likewise, the addition of a 4-person family to a neighbourhood is expected to increase the neighbourhood’s average after-tax income by approximately \$148.40 (± 16.32). Interestingly, the addition of a 3-person census family to a neighbourhood is expected to decrease the neighbourhood’s average after-tax income by around \$184.42 (± 17.66). These findings seem to suggest that 4-person families are also be key drivers of average neighbourhood income. Additionally, this may imply that on average, while 2 and 4-person families can commonly have above-average incomes, more 3-person families may have below-average incomes across Toronto’s neighbourhoods.

## Map Visualization

![Distribution of 4-person Census Families Across Toronto](images/map_visualization_four-02.png){#fig-map}

![Distribution of 2-person Census Families Across Toronto](images/map_visualization_two-02.png){#fig-map2}

The initial analysis of the distribution of different census family sizes across Toronto neighbourhoods highlighted that 2-person families can be key drivers of average neighbourhood income. The results of the model further suggest that 2 and 4-person families positively contribute to average neighbourhood income. With this in mind, @fig-map maps the distribution of 4-person census families throughout the city of Toronto. This shows how the majority of 4-person census families are found in the neighbourhoods towards the east end of the city with a few neighbourhoods near the city’s centre and the Harbourfront area. @fig-map2 demonstrates the distribution of 2-person families. Here, there appears to be a shift in that more 2-person census families seem to be in neighbourhoods closer to the city’s centre and the west side of the Harbourfront area. These illustrations provide a better understanding of where the key drivers of average neighbourhood income are located.

@@

The construction of a model for this analysis began by fitting an initial model with average neighbourhood income as the response variable and 2, 3, 4, and 5 or more-person families as predictors. As seen by @tbl-initialmodel and the individual t-tests, 5 or more-person families do not appear to significantly impact a neighbourhood’s average income with the largest p-value among all predictors (p \> 0.001).

@@

So, another model was created without this variable (i.e. a reduced model), and all predictors seem to have significant impacts (p \< 0.001) on neighbourhood income as shown in @tbl-newmodel.

@@

par(mfrow=c(1,3))
plot(residuals_1 ~ regression_data$size_two, main = "Residuals vs 2-person Families", 
     xlab = "Number of 2-person Families", ylab = "Residuals")
plot(residuals_1 ~ regression_data$size_three, main = "Residuals vs 3-person Families", 
     xlab = "Number of 3-person Families", ylab = "Residuals")
plot(residuals_1 ~ regression_data$size_four, main = "Residuals vs 4-person Families", 
     xlab = "Number of 4-person Families", ylab = "Residuals")
```

The Residual vs Fitted Values plot (@fig-residualsandfitted) seems to have some clustering of points, indicating that the constant variance and independence assumptions are somewhat satisfied. However, because the overall pattern of the plot is still relatively spread out with no observable patterns, and as the objective of the model is interpretability, transformations were not applied to the variables in the model to better these assumptions.

@@

plot(residuals_1 ~ fitted_1, main = "Residuals vs Fitted Values", 
     xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col ="red")
```

Lastly, though the Normal Q-Q plot (@fig-normal) shows some deviation from the Q-Q line, the normality assumption is somewhat satisfied with little evidence to suggest an alarming violation of this assumption. Putting all this together, while improvements can be made to the model for better, more accurate predictions, it is sufficient for the purpose of interpretability.
